The Increased Effect of Elections and Changing Prime Ministers on Topics Discussed in the Australian Federal Parliament between 1901 and 2018

### 2019-10-30 outputs/docs/paper/paper_MA_structure_edit.Rmd
intro edits

---
output:
  bookdown::pdf_document2:
    citation_package: natbib
    keep_tex: false
    fig_caption: true
    number_sections: true
    latex_engine: pdflatex
    template: svm-latex-ms.tex
title: "The Increased Effect of Elections and Prime Ministers on Discussion in the Australian Federal Parliament between 1901 and 2018"
thanks: "Thank you to Chris Cochrane, Dan Simpson, Jill Sheppard, John McAndrews, John Tang, Leslie Root, Martine Mariotti, Matt Jacob, Matthew Kerby, Myles Clark, Ruth Howlett, Tianyi Wang, Tim Hatton, and Zach Ward for their helpful suggestions; and to the UC Berkeley Demography Department for the use of their computing resources. We are grateful for the many excellent comments that we received from seminar participants at the ANU SPIR, the ANU RSE, the Australian Parliamentary Library, the Max Planck Institute for Demographic Research, the U of T Political Behavior Group, and the 2019 PSA Political Methodology Conference. The code and data for this paper is available at: https://github.com/RohanAlexander/hansard. Comments and suggestions on the `r format(Sys.time(), '%d %B %Y')` version of this paper welcome at: rohan.alexander@anu.edu.au."
author:
- name: Rohan Alexander
  affiliation: Australian National University
- name: Monica Alexander
  affiliation: University of Toronto
abstract: "A large part of politics consists of reacting to events, but not all events provoke a reaction. As such, understanding political behaviour requires examining the reaction to events and how this changes over time. We systematically analyse how discussion in the Australian Federal Parliament changes in response to two types of events: elections and changed prime ministers. To do this we first create a new dataset of what was said in the Australian Federal Parliament from 1901 through to 2018 based on available public records. We reduce the dimensionality of this dataset using a correlated topic model, relate those topics to the Comparative Agendas Project, and then analyse the effect of these two types of events using a Bayesian hierarchical Dirichlet model that has several advantages over existing models. We find that: changes in prime minister tend to be associated with topic changes even when the party in power does not change; and elections that do not result in a change in prime minister are rarely associated with topic changes in the first half of our sample, but have been increasingly significant since the 1980s."
keywords: "text-as-data, Australian politics, unsupervised machine learning, Bayesian hierarchical Dirichlet model, Comparative Agendas Project"
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 12pt
toc: FALSE
# spacing: double
bibliography: ../bibliography.bib
biblio-style: apsr
endnote: no

@@

Events are a well-known cause of political behaviour. But their effect is varied, and changes over time. Here we focus on just two types of events: elections, and changes in prime minister. We examine the effect of these types of events on the topics that are discussed in the Australian Federal Parliament.

The main data source for our paper is the text record, known as 'Hansard', of what was said in the Australian Federal Parliament. From this we create a dataset based on the sitting days from 1901 through to 2018. We have records for 7,934 days in the House of Representatives (lower house) and 6,746 days in the Senate (upper house). In contrast to the US, but similar to Canada and the UK, in Australia the executive is drawn from the legislature, and the prime minister typically sits in the lower house. Unlike Canada or the UK, both houses are elected, although each has different mechanisms and constituencies. Our dataset consists of counts of every word that occurs at least thrice, after we convert all words to lower-case; remove names, numbers, and punctuation; join commonly co-located words; and remove stop words.

We systematically analyse the text in two stages, each of which takes advantage of a different statistical technique. First, to reduce the dimensionality of our dataset we use a topic model that allows correlation between topics, and then further reduce those topics to match the Comparative Agendas Project. Second, we analyse those topics using a Bayesian hierarchical Dirichlet model to examine changes at various types of events. We explicitly consider two types of events: changes in prime minister and elections.

@@

Our work contributes to the growing literature that analyses text using quantitative methods. It sits across, and draws from, various historically-separate disciplines including applied statistics, economics, and political science. In addition to our findings about the effect of events in Australian politics, we contribute to this literature in terms of both data and methods. From a data perspective we bring to bear an essentially-complete record of what was said in the Australian Federal Parliament on a daily basis, and our dataset is available to other researchers. From a methods perspective, our analysis model has several advantages over existing methods. These include: allowing the events to have more-complicated auto-correlated functional forms; implementing pooling across groups of similar documents; and identifying outlying topic distributions without the need to pre-specify the event of interest. There are many avenues for closely related future work such as: investigating the effect of other types of events; including a richer set of covariates to disentangle the reason for different effects; and reversing the causality to examine the effect of what is said in parliament on various outcomes.

>>>>

---
output:
  bookdown::pdf_document2:
    citation_package: natbib
    keep_tex: false
    fig_caption: true
    number_sections: true
    latex_engine: pdflatex
    template: svm-latex-ms.tex
title: "The Increased Effect of Elections and Changing Prime Ministers on Topics Discussed in the Australian Federal Parliament between 1901 and 2018"
thanks: "Thank you to Chris Cochrane, Dan Simpson, Jill Sheppard, John McAndrews, John Tang, Leslie Root, Martine Mariotti, Matt Jacob, Matthew Kerby, Myles Clark, Ruth Howlett, Tianyi Wang, Tim Hatton, and Zach Ward for their helpful suggestions; and to the UC Berkeley Demography Department for the use of their computing resources. We are grateful for the many excellent comments that we received from seminar participants at the ANU SPIR, the ANU RSE, the Australian Parliamentary Library, the Max Planck Institute for Demographic Research, the U of T Political Behavior Group, and the 2019 PSA Political Methodology Conference. The code and data for this paper is available at: https://github.com/RohanAlexander/hansard. Comments and suggestions on the `r format(Sys.time(), '%d %B %Y')` version of this paper welcome at: rohan.alexander@anu.edu.au."
author:
- name: Rohan Alexander
  affiliation: Australian National University
- name: Monica Alexander
  affiliation: University of Toronto
abstract: "Politics and discussion in parliament is likely to be influenced by the party in power and associated election cycles. However, little is known about the extent to which these events affect discussion and how this has changed over time. We systematically analyse how discussion in the Australian Federal Parliament changes in response to two types of political events: elections and changed prime ministers. We use a newly constructed dataset of what was said in the Australian Federal Parliament from 1901 through to 2018 based on extracting and cleaning available public records. We reduce the dimensionality of discussion in this dataset by using a correlated topic model to obtain a set of comparable topics over time. We then relate those topics to the Comparative Agendas Project, and then analyse the effect of these two types of events using a Bayesian hierarchical Dirichlet model. We find that: changes in prime minister tend to be associated with topic changes even when the party in power does not change; and the effect of elections has been increasing since the 1980s, regardless of whether the election results in a change of prime minister."
keywords: "text-as-data, Australian politics, unsupervised machine learning, Bayesian hierarchical Dirichlet model, Comparative Agendas Project"
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 12pt
toc: FALSE
# spacing: double
bibliography: ../bibliography.bib
biblio-style: apsr
endnote: no

@@

What is discussed in parliament is a key indicator of a government's priorities and likely policy outcomes. While in an ideal world, we may hope that topics of discussion represent the most important social, economic and environmental issues for the country at that particular time, in reality, discussion is heavily influenced by party politics, the priorities of the head of government at the time, and where they are in the current election cycle.

For example, XX NEED A BIT OF A LITERATURE REVIEW HERE on papers that look at discussion and political changes?? XX

While there is evidence to suggest a relationship between parliamentary discussion and political events, little research exists that systematically analyses the sensitivity of topics discussed in parliament to political changes in a statistical way. This is mostly likely because historical data are difficult to obtain, and parliamentary discourse and events are inherently complex and multi-dimensional, and difficult to operationalise into an analysable form. However, the increased accessibility of optical character recognition to parse historical documents, and recent developement in statistical text analysis, allow for a more broad analysis of parliamentary events and discussion over time.

In this paper, we exame the effect of elections and changes in prime ministers on the topics that are discussed in the Australian Federal Parliament. We construct a new dataset from the text record, known as 'Hansard', of what was said in the Australian Federal Parliament since its inception in 1908, to 2018. The dataset covers records for 7,934 days in the House of Representatives (lower house) and 6,746 days in the Senate (upper house).

We then systematically analyse the text contained in the Hansard dataset in two stages, each of which takes advantage of a different statistical technique. First, we used a correlated topic model to obtain a set of comparable topics discussed in parliament over time, and then further reduce those topics to match the Comparative Agendas Project. Second, we introduce a new modeling approach to examine the association between changes in topics discussed changes in prime minister and elections. The second modeling approach centers on a Bayesian hierarchical Dirichlet model, which allows for the effect of these two event types to be modeled concurrently, while taking time since election into account. The model also allows for information about event effects to be pooled across documents.

@@

Our work contributes to the growing literature that analyses text using quantitative methods. It sits across, and draws from, various historically-separate disciplines including applied statistics, economics, and political science. In addition to our findings about the effect of events in Australian politics, we contribute to this literature in terms of both data and methods. From a data perspective we bring to bear an essentially-complete record of what was said in the Australian Federal Parliament on a daily basis, and our dataset is available to other researchers. From a methods perspective, we present a modeling framework that can be generalizable to other problems looking at the association between text data and events. The model allows for time since event to be taken into account; implements pooling across groups of similar documents; and additionally identifies potential outlying topic distributions without the need to pre-specify the event of interest. There are many avenues for closely related future work such as: investigating the effect of other types of events; including a richer set of covariates to disentangle the reason for different effects; and reversing the causality to examine the effect of what is said in parliament on various outcomes.

The remainder of the paper is structured as follows XX
